What defines big team science?

### 2025-06-10 manuscript_definition.Rmd
edits edits edits

  Funding: No funding was obtained for this project.

@@

  Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun). After review, we updated the manuscript to reflect that the definition of big team science is unclear, and focused on a data-driven definition to explore our research questions.

@@

keywords          : "big team, science, authorship, credit"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
classoption       : "man"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib

@@

Scientific discovery has increasingly become a collaborative process,
and the scale and scope of team science have dramatically expanded in
recent years [@council2015]. Collaboration in scientific endeavors
involves multiple researchers at (potentially) multiple institutions to
communicate and work together to advance knowledge in their chosen
field(s). Collaboration can manifest uniquely in each project dependent
on the skill sets, hypotheses, and perspectives of collaborators. One
key aspect of collaboration is the flexibility allotted as it is shaped
by the needs of the project and the researchers involved. While
collaboration is not new in science, interest in "team science" is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Team science is
often defined as groups of researchers with various expertise working
together to investigate complex problems [@fiore2008]. Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently.

The movement toward team science reflects demands of modern research to
answer complex questions, meet funder and university desires for
interdisciplinary research, and the desire to increase diversity of
scientific representation [@council2015]. Further, the evolution of team
science reflects broader shifts in research practices, driven by two
sources: 1) increasing globalization and technology that allows for
real-time interdisciplinary research [@jones2008], and 2) expanding
interest in reproducibility, replication, and generalizability [i.e.,
the credibility movement, @maxwell2015; @nelson2018; @zwaan2018;
@vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

@@

Our findings build on previous research by also examining diversity in
author seniority and geopolitical affiliation. The growing participation
of early-career scholars over time suggests that big team science may be
increasingly accessible to a broader range of researchers, not just
senior or established scientists. This trend is interesting given the
challenges BTS projects can pose for non-permanent researchers: slow
publication timelines, uncertain publication outcomes, and fewer
incentives for non-corresponding authors. Yet, large teams may allow for
more distributed workloads and reduced individual time investment, which
could make them appealing even for early-career researchers. Moreover,
prior work has shown that publications from larger teams tend to receive
more citations and have broader impact [[@larivière2015], which may
further incentivize early-career involvement despite the structural
risks. Globalization, the internet, and the focus on interdisciplinary
research are potentially driving forces behind our results, but,
hopefully, the results also point to a decline in scientific gatekeeping
[@lu2007; @siler2015].

>>>>

  Funding: No funding was obtained for this project. #we staying with this even though it helped with coding errors?

@@

  Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun). After peer review, the preregisteration was updated to address the unclear definition of big team science to focus on a data-driven definition to explore the research questions.

@@

keywords          : "big team science, team science, authorship, credit"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
classoption       : "man"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib

@@

Scientific discovery has increasingly become a collaborative process,
with the scale and scope of team science have dramatically expanded in
recent years [@council2015]. Collaboration in scientific endeavors
involves multiple researchers (potentiality) at multiple institutions to
communicate and work together to advance knowledge in their chosen
field(s). This unique composition of a collaboration for each project is 
dependent on the skill sets, hypotheses, and perspectives of collaborators 
involved. A key strength of collaboration lies in its flexibility, allowing it 
to adapt to the needs of the project and the researchers themselves.

While collaboration is not new in science, interest in "team science" is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Team science is
often defined as groups of researchers with various expertise working
together to investigate complex problems [@fiore2008]. Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently.

The movement toward team science reflects demands of modern research to
answer complex questions, meet funding agencies and universities desires for
interdisciplinary research, and the desire to increase scientific representation
[@council2015]. Further, the evolution of team science reflects broader shifts
in research practices, driven by two sources: 1) increasing globalization 
and technology that allows for real-time interdisciplinary research [
@jones2008], and 2) expanding interest in reproducibility, replication, 
and generalizability [i.e., the credibility movement, @maxwell2015; @nelson2018; 
@zwaan2018; @vazire2022]. Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

@@

Our findings build on previous research by also examining diversity in
author seniority and geopolitical affiliation. The growing participation
of early-career scholars over time suggests that big team science may be
increasingly accessible to a broader range of researchers, not just
senior or established scientists. This trend is interesting given the
challenges BTS projects can pose for non-permanent researchers: slow
publication timelines, uncertain publication outcomes, and fewer
incentives for non-corresponding authors. Yet, large teams may allow for
more distributed workloads and reduced individual time investment, which
could make them appealing even for early-career researchers. Moreover,
prior work has shown that publications from larger teams tend to receive
more citations and have broader impact [@larivière2015], which may
further incentivize early-career involvement despite the structural
risks. Globalization, the internet, and the focus on interdisciplinary
research are potentially driving forces behind our results, but,
hopefully, the results also point to a decline in scientific gatekeeping
[@lu2007; @siler2015].

### 2025-06-10 manuscript_definition.tex
edits edits edits

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{big team, science, authorship, credit}
\usepackage{lineno}

@@

\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={What defines big team science?},
  pdfauthor={Erin M. Buchanan1 \& Savannah C. Lewis2},
  pdflang={en-EN},
  pdfkeywords={big team, science, authorship, credit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Funding: No funding was obtained for this project.

@@

Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (\url{https://osf.io/f2dtr}) but then was updated with access to the Scopus database for a broader picture of BTS projects (\url{https://osf.io/fheun}). After review, we updated the manuscript to reflect that the definition of big team science is unclear, and focused on a data-driven definition to explore our research questions.

@@

Scientific discovery has increasingly become a collaborative process,
and the scale and scope of team science have dramatically expanded in
recent years (Council et al., 2015). Collaboration in scientific endeavors
involves multiple researchers at (potentially) multiple institutions to
communicate and work together to advance knowledge in their chosen
field(s). Collaboration can manifest uniquely in each project dependent
on the skill sets, hypotheses, and perspectives of collaborators. One
key aspect of collaboration is the flexibility allotted as it is shaped
by the needs of the project and the researchers involved. While
collaboration is not new in science, interest in ``team science'' is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Team science is
often defined as groups of researchers with various expertise working
together to investigate complex problems (Fiore, 2008). Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently.

The movement toward team science reflects demands of modern research to
answer complex questions, meet funder and university desires for
interdisciplinary research, and the desire to increase diversity of
scientific representation (Council et al., 2015). Further, the evolution of team
science reflects broader shifts in research practices, driven by two
sources: 1) increasing globalization and technology that allows for
real-time interdisciplinary research (B. F. Jones et al., 2008), and 2) expanding
interest in reproducibility, replication, and generalizability (i.e.,
the credibility movement, Maxwell et al., 2015; Nelson et al., 2018; Vazire et al., 2022; Zwaan et al., 2018). Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

@@

Our findings build on previous research by also examining diversity in
author seniority and geopolitical affiliation. The growing participation
of early-career scholars over time suggests that big team science may be
increasingly accessible to a broader range of researchers, not just
senior or established scientists. This trend is interesting given the
challenges BTS projects can pose for non-permanent researchers: slow
publication timelines, uncertain publication outcomes, and fewer
incentives for non-corresponding authors. Yet, large teams may allow for
more distributed workloads and reduced individual time investment, which
could make them appealing even for early-career researchers. Moreover,
prior work has shown that publications from larger teams tend to receive
more citations and have broader impact {[}(Larivière et al., 2015), which may
further incentivize early-career involvement despite the structural
risks. Globalization, the internet, and the focus on interdisciplinary
research are potentially driving forces behind our results, but,
hopefully, the results also point to a decline in scientific gatekeeping
(Lu, 2007; Siler et al., 2015).

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\makeatother
\keywords{big team science, team science, authorship, credit}
\usepackage{lineno}

@@

\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={What defines big team science?},
  pdfauthor={Erin M. Buchanan1 \& Savannah C. Lewis2},
  pdflang={en-EN},
  pdfkeywords={big team science, team science, authorship, credit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Funding: No funding was obtained for this project. \#we staying with this even though it helped with coding errors?

@@

Pre-registration: This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (\url{https://osf.io/f2dtr}) but then was updated with access to the Scopus database for a broader picture of BTS projects (\url{https://osf.io/fheun}). After peer review, the preregisteration was updated to address the unclear definition of big team science to focus on a data-driven definition to explore the research questions.

@@

Scientific discovery has increasingly become a collaborative process,
with the scale and scope of team science have dramatically expanded in
recent years (Council et al., 2015). Collaboration in scientific endeavors
involves multiple researchers (potentiality) at multiple institutions to
communicate and work together to advance knowledge in their chosen
field(s). This unique composition of a collaboration for each project is
dependent on the skill sets, hypotheses, and perspectives of collaborators
involved. A key strength of collaboration lies in its flexibility, allowing it
to adapt to the needs of the project and the researchers themselves.

While collaboration is not new in science, interest in ``team science'' is
growing as individual researchers seek an interdisciplinary approach to
research or bring on more students to their project. Team science is
often defined as groups of researchers with various expertise working
together to investigate complex problems (Fiore, 2008). Unlike general
collaboration, team science involves structured roles, coordinated
workflows, and shared resources to address challenges that would be
difficult for individual researchers or one small team to solve
independently.

The movement toward team science reflects demands of modern research to
answer complex questions, meet funding agencies and universities desires for
interdisciplinary research, and the desire to increase scientific representation
(Council et al., 2015). Further, the evolution of team science reflects broader shifts
in research practices, driven by two sources: 1) increasing globalization
and technology that allows for real-time interdisciplinary research (B. F. Jones et al., 2008), and 2) expanding interest in reproducibility, replication,
and generalizability (i.e., the credibility movement, Maxwell et al., 2015; Nelson et al., 2018; Vazire et al., 2022; Zwaan et al., 2018). Technological advances have provided easier ways to
collaborate with people who are from other universities and countries
through document sharing platforms (e.g., Google, GitHub, and the Open
Science Framework), video chatting platforms (e.g., Zoom, Microsoft
Teams), and messaging and project management platforms (e.g., Slack,
Trello, when2meet, etc.).

@@

Our findings build on previous research by also examining diversity in
author seniority and geopolitical affiliation. The growing participation
of early-career scholars over time suggests that big team science may be
increasingly accessible to a broader range of researchers, not just
senior or established scientists. This trend is interesting given the
challenges BTS projects can pose for non-permanent researchers: slow
publication timelines, uncertain publication outcomes, and fewer
incentives for non-corresponding authors. Yet, large teams may allow for
more distributed workloads and reduced individual time investment, which
could make them appealing even for early-career researchers. Moreover,
prior work has shown that publications from larger teams tend to receive
more citations and have broader impact (Larivière et al., 2015), which may
further incentivize early-career involvement despite the structural
risks. Globalization, the internet, and the focus on interdisciplinary
research are potentially driving forces behind our results, but,
hopefully, the results also point to a decline in scientific gatekeeping
(Lu, 2007; Siler et al., 2015).
